Sparse Word Embeddings for Multi-domain Machine Translation

### 2019-05-16 2019/emnlp/sparse.tex
add / remove comments

We experiment with two language pairs (English-French, English-German )and data originating from three domains, corresponding to texts from three European institutions: from the European Parlement (EPPS) \cite{Koehn05europarl}, the European Medicines Agency (EMEA), and the European Central Bank (ECB) \cite{Tiedemann2009RANLP5}. We randomly split those corpora into training, validation and test sets (see statistics in Table~\ref{tab:Corpora}). The validation sets are used to chose the best model according to the average BLEU score. Using internal test sets provides us with somewhat ideal conditions; for contrast experiments, we also use supplementary test sets from other, related, domains: the official Khresmoi-test \fyTodo{REF needed}, which is close to EMEA, and News test 2014 \fyTodo{REF needed}. This enables us to evaluate the loss in performance when the test set is from a different domain.
% The model is also required to achieve comparable performance to generic model. To do so, we use newstest 2009 and IWSLT 2010 whose contain does not particularly belong to any domain.
\fyTodo{Check which corpus are useful}
\begin{table}
  \centering
  \begin{tabular}{ *{4}{|r|}}
    \hline
    Corpus & Train & Dev & Test \\ \hline
    \multicolumn{4}{|l|}{$En \rightarrow Fr$ }\\
    \multicolumn{4}{|l|}{Vocab: En 30,165, Fr: 30,398}\\
    \hline
    EMEA  & 1,09M & 1000 & 1000 \\
    ECB    & 0,19M & 1000 & 1000     \\
    EPPS   & 2,0M  & 1000 & 1000  \\ \hline \hline
    \multicolumn{4}{|l|}{$En \rightarrow Ge$}\\
    \multicolumn{4}{|l|}{ Vocab: En 30,159, Ge: 30,698}\\ \hline
    EMEA  & 1,11M & 1000 & 1000 \\
    ECB     &  0,11M & 1000 & 1000  \\
    EPPS   & 1,92M & 1000 & 1000 \\ \hline
\end{tabular}
\caption{Corpora}
\label{tab:Corpora}
\end{table}

To reduce the vocabulary, we apply Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations, then select approximately 30,000 the most frequent words to build vocabulary for each language.\fyTodo{I need explanations here}

@@

For all models, we set the size of word embeddings equal to 512; the size of hidden layers is equal to 1024 for RNNs and 512 for Transformer.
\fyTodo{Improve this}
\begin{itemize}
\item Transformer: size of word embedding: 512; size of hidden layers: 512; size of inner feed forward layer: 2048; Multi-head: 8; number of layers: 6.
\item Attention Recurrent Neural Network: Bidirection LSTM encoder, unidirectional LSTM decoder, Luong Attention Mechanism, size of word embedding: 512; size of hidden layers: 1024; number of layers: 1
\item DC: size of domain embedding:2 ; same settings for both architecture: Transformer and Attention Recurrent Neural Network.
\item WDCMT: Bidirection Gated recurrent units (denoted GRU) encoder, unidirectional conditional GRU decoder, size of word embedding: 512; size of hidden layers: 1024; size of domain layers: 300; number of layers: 1
\end{itemize}
To train NMT systems, we use Adam($\beta_1=0.9$, $\beta_2 = 0.999$, $\alpha=0.0005$) for RNNs; and Adam with $\beta_1=0.9$, $\beta_2= 0.98$, with Noam decay \cite{Vaswani17attention} ($warmup\_steps=4000$) for Transformer.
Our implementation is from OpenNMT-tf\cite{Klein2017OpenNMT}, except for WDCMT where we use the authors implementation\footnote{\url{https://github.com/DeepLearnXMU/WDCNMT}.}.

\subsection{Sparse NMT model configuration}
\fyTodo{Motivate the split - discuss experimentally embedding size}
For the implementation of sparse embeddings, we split the vector embedding of size 512 into four regions: 3 domain specific and 1 generic, with size $[8,8,8,488]$, the choice will be dicussed in \ref{secc:region_size}. If sentence comes from domain $i$, regions for domains $j \neq i$ will be masked as 0 while generic region is always activated. We then use a dense layer of size 512 to fuse the active domain region and the generic region.

@@

\subsection{Results}
\begin{table*}
\begin{center}
 \begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
 \hline
 Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
 \hline\hline
 Mixed & 67.69 & 37.5 & 53.49 & 34.86 & 30.96 & 20.98 & 28.56 & 25.7 \\
 \hline
 EMEA-finetuned & 76.77 & 17.16 & 11.99 & 29.58 & 14.25 & 9.92 & 11.64 & 11.1 \\
 \hline
 EPPS-finetuned & 20.86 & 37.04 & 24.53 & 31.86 & 30.97 & 20.72 & 28.33 & 11.1 \\
 \hline
 ECB-finetuned & 26.93 & 27.09 & 56.52 & 22.04 & 22.92 & 12.94 & 17.51 & 13.99 \\
 \hline
 Sparse (1) & 74.26 & 37.67 & 54.07 & 33.85 & 30.86 & 20.97 & 28.11 & 25.7 \\
 \hline
 Sparse (2) & 73.97 & 37.81 & 53.67 & 34.17 & 30.93 & 20.97 & 28.11 & 25.7 \\
 \hline
 Sparse (3) &  &  &  &  &  &  &  & \\
 \hline
 Domain Control & 67.87 & 37.31 & 54.14 & 33.63 & 31.09 &  20.1 & 26.52 & 24.81\\
 \hline
\end{tabularx}
\end{center}

\caption{BPE-detokenized BLEU score for English-French with the Transformer architecture. Sparse(1) uses correponding domain tag for EMEA, EPPS, ECB test sets and generic tag for newstest and IWSLT. Sparse(2) uses generic tag for all of test sets. Sparse(3) uses tag predicted by a pre-trained domain classififer}
\label{tab:1}
\end{table*}

@@

\subsection{Region's size \label{secc:region_size}}
We also test different setting of region size in word embedding. However, we limit our research in 4 settings: $[4,4,4,496]$, $[8,8,8,488]$, $[16,16,16,464]$, $[32,32,32,32,416]$. We test those settings with the same condition as above. The scores is showed in table \ref{tab:5}
\begin{table*}
\begin{center}
 \begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X | X ||} 
 \hline
 Models & EMEA test & EMEA excluded & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
 \hline\hline 
 Sparse - 4 & 74.65 & 49.61 & 37.42 & 54.49 & 33.68 & 31.01 & 20.71 & 28.03 & 25.28 \\
 \hline
 Sparse - 8 & 74.26 & 49.9 & 37.67 & 54.07 & 33.85 & 30.86 & 20.97 & 28.11 & 25.7 \\
 \hline
 Sparse - 16 & 74.15 & 49.1 & 37.78 & 54.56 & 33.4 & 30.93 & 20.73 & 28.1 & 25.33 \\
 \hline
 Sparse - 32 & 75.1 & 48.61 & 37.64 & 54.29 & 33.2 & 30.94 & 21.07 & 27.99 & 25.44 \\
 \hline
\end{tabularx}
\end{center}
\caption{BPE-detokenized BLEU score for English-French with the Transformer architecture}
\label{tab:5}
\end{table*}

@@

\begin{table}
  \centering
  \begin{tabular}{*{4}{|r|}}
    \hline
    Domain & Word & Neighbor & Cosinus distance \\ \hline
    \multirow{5}{*}{EMEA} & \multirow{5}{*}{joint} 
    & Joint & 0.371 \\  
    &  & jointly & 0.545 \\
    &  & artic￭ & 0.605 \\
    &  & joints & 0.607 \\
	&  & collective & 0.609 \\
	\hline
	\multirow{5}{*}{EPPS} & \multirow{5}{*}{joint} 
	& Joint & 0.365 \\
	& & jointly & 0.497 \\
	& & join￭ & 0.606 \\
	& & common & 0.607 \\
	& & collective & 0.626 \\
	& & artic￭ & 0.633\\
	\hline
	\multirow{5}{*}{ECB} & \multirow{5}{*}{joint}
	& Joint & 0.423 \\
	& & jointly & 0.467 \\
	& & common & 0.601 \\
	& & together & 0.632 \\
	& & artic￭ & 0.663 \\
	& & coordinated & 0.668 \\
\end{tabular}
\caption{Corpora}
\label{tab:Polysemy}
\end{table}

>>>>

We experiment with two language pairs (English-French, English-German) and data originating from three domains, corresponding to texts from three European institutions: from the European Parlement (EPPS) \cite{Koehn05europarl}, the European Medicines Agency (EMEA), and the European Central Bank (ECB) \cite{Tiedemann2009RANLP5}. We randomly split those corpora into training, validation and test sets (see statistics in Table~\ref{tab:Corpora}). The validation sets are used to chose the best model according to the average BLEU score. Using internal test sets provides us with somewhat ideal conditions; for contrast experiments, we also use supplementary test sets from other related domains: the official Khresmoi-test \cite{Khresmoi17test}, which is close to EMEA, News test 2014 \cite{Bojar14findings}, IWSLT 2010 (Ted Talks) \cite{}. This enables us to evaluate the loss in performance when the test set is from a different domain.
% The model is also required to achieve comparable performance to generic model. To do so, we use newstest 2009 and IWSLT 2010 whose contain does not particularly belong to any domain.
\fyDone{Check which corpus are useful}
\begin{table}
  \centering
  \begin{tabular}{ *{4}{|r|}}
    \hline
    Corpus & Train & Dev & Test \\ \hline
    \multicolumn{4}{|l|}{$En \rightarrow Fr$ }\\
    \multicolumn{4}{|l|}{Vocab: En 30,165, Fr: 30,398}\\
    \hline
    EMEA  & 1,09M & 1000 & 1000 \\
    ECB    & 0,19M & 1000 & 1000     \\
    EPPS   & 2,0M  & 1000 & 1000  \\ \hline \hline
    \multicolumn{4}{|l|}{$En \rightarrow Ge$}\\
    \multicolumn{4}{|l|}{ Vocab: En 30,159, Ge: 30,698}\\ \hline
    EMEA  & 1,11M & 1000 & 1000 \\
    ECB     &  0,11M & 1000 & 1000  \\
    EPPS   & 1,92M & 1000 & 1000 \\ \hline
\end{tabular}
\caption{Corpora}
\label{tab:Corpora}
\end{table}

To reduce the vocabulary, we apply Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations separately for each language. \fyDone{I need explanations here}

@@

For all models, we set the size of word embeddings equal to 512; the size of hidden layers is equal to 1024 for RNNs and 512 for Transformer.
\fyTodo{Improve this}
\begin{itemize}
\item Transformer: size of word embedding: 512; size of hidden layers: 512; size of inner feed forward layer: 2048; Multi-head: 8; number of layers: 6.
\item Attention Recurrent Neural Network: Bidirection LSTM encoder, unidirectional LSTM decoder, Luong Attention Mechanism, size of word embedding: 512; size of hidden layers: 1024; number of layers: 1
\item DC: size of domain embedding:2 ; same settings for both architecture: Transformer and Attention Recurrent Neural Network.
\item WDCMT: Bidirection Gated recurrent units (denoted GRU) encoder, unidirectional conditional GRU decoder, size of word embedding: 512; size of hidden layers: 1024; size of domain layers: 300; number of layers: 1
\end{itemize}
To train NMT systems, we use Adam($\beta_1=0.9$, $\beta_2 = 0.999$, $\alpha=0.0005$) for RNNs; and Adam with $\beta_1=0.9$, $\beta_2= 0.98$, with Noam decay \cite{Vaswani17attention} ($warmup\_steps=4000$) for Transformer.
Our implementation is from OpenNMT-tf\cite{Klein2017OpenNMT}, except for WDCMT where we use the authors implementation.\footnote{\url{https://github.com/DeepLearnXMU/WDCNMT}.}

\subsection{Sparse NMT model configuration}\fyTodo{Change sparse.}
\fyTodo{Motivate the split - discuss experimentally embedding size}
For the implementation of sparse embeddings, we split the vector embedding of size 512 into four regions: 3 domain specific and 1 generic, with size $[8,8,8,488]$, the choice will be dicussed in \ref{secc:region_size}. If sentence comes from domain $i$, regions for domains $j \neq i$ will be masked as 0 while generic region is always activated. We then use a dense layer of size 512 to fuse the active domain region and the generic region.

@@

\subsection{Results}
Discussion 1:
\begin{itemize}
\item transformer better than RNN
\item Fine tuning is the best strategy - but subject to catastrophic
\item Sparse better than joint - better than domain control
\item Across th board / language \& architecture - always good for EMEA , less good for EPPS, varying for ECB
\item better ? than domain control ? WMCD ?
\end{itemize}

Discussion 2:
\begin{itemize}
\item Sparse 1 > Sparse 2
\item out of domain better than Mixed
\item Sparse 3 ??
\item ajouter Sparse 4
\end{itemize}

\begin{table*}
\begin{center}
 \begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
 \hline
 Models & EMEA & EPPS & ECB & Khresmoi & test 2007 (EPPS domain) & newstest 2009 & Newstest 2014 & IWSLT 2010 \\ [0.5ex] 
 \hline\hline
 Mixed & 67.69 & 37.5 & 53.49 & 34.86 & 30.96 & 20.98 & 28.56 & 25.7 \\
 \hline
 EMEA-finetuned & 76.77 & 17.16 & 11.99 & 29.58 & 14.25 & 9.92 & 11.64 & 11.1 \\
 \hline
 EPPS-finetuned & 20.86 & 37.04 & 24.53 & 31.86 & 30.97 & 20.72 & 28.33 & 11.1 \\
 \hline
 ECB-finetuned & 26.93 & 27.09 & 56.52 & 22.04 & 22.92 & 12.94 & 17.51 & 13.99 \\
 \hline
 Sparse (1) & 74.26 & 37.67 & 54.07 & 33.85 & 30.86 & 20.97 & 28.11 & 25.7 \\
 \hline
 Sparse (2) & 73.97 & 37.81 & 53.67 & 34.17 & 30.93 & 20.97 & 28.11 & 25.7 \\
 \hline
 Sparse (3) &  &  &  &  &  &  &  & \\
 \hline
 Domain Control & 67.87 & 37.31 & 54.14 & 33.63 & 31.09 &  20.1 & 26.52 & 24.81\\
 \hline
\end{tabularx}
\end{center}

\caption{BPE-detokenized BLEU score for English-French with the Transformer architecture. Sparse(1) uses correponding domain tag for EMEA, EPPS, ECB test sets and generic tag for newstest and IWSLT. Sparse (2) uses generic tag for all of test sets. Sparse (3) uses tag predicted by a pre-trained domain classififer}
\label{tab:1}
\end{table*}

@@

\subsection{Region's size \label{secc:region_size}}

We also test different setting of region size in word embedding. However, we limit our research in 4 settings: $[4,4,4,496]$, $[8,8,8,488]$, $[16,16,16,464]$, $[32,32,32,32,416]$. We test those settings with the same condition as above. The scores are in Table~\ref{tab:embedding-size}.\fyTodo{Comment, homegenize captions.}

Results are very close so we pick 8.\fyTodo{Experiment with 64}.

\begin{table*}
\begin{center}
 \begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X | X ||} 
 \hline
 Models & EMEA test & EMEA excluded & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
 \hline\hline 
 Sparse - 4 & 74.65 & 49.61 & 37.42 & 54.49 & 33.68 & 31.01 & 20.71 & 28.03 & 25.28 \\
 \hline
 Sparse - 8 & 74.26 & 49.9 & 37.67 & 54.07 & 33.85 & 30.86 & 20.97 & 28.11 & 25.7 \\
 \hline
 Sparse - 16 & 74.15 & 49.1 & 37.78 & 54.56 & 33.4 & 30.93 & 20.73 & 28.1 & 25.33 \\
 \hline
 Sparse - 32 & 75.1 & 48.61 & 37.64 & 54.29 & 33.2 & 30.94 & 21.07 & 27.99 & 25.44 \\
 \hline
\end{tabularx}
\end{center}
\caption{BPE-detokenized BLEU score for English-French with the Transformer architecture}
\label{tab:5}
\end{table*}

@@

\begin{table}
  \centering
  \begin{tabular}{*{4}{|r|}}
    \hline
    Domain & Word & Neighbor & Cosinus distance \\ \hline
    \multirow{5}{*}{EMEA} & \multirow{5}{*}{joint} 
    & Joint & 0.371 \\  
    &  & jointly & 0.545 \\
    &  & artic- & 0.605 \\
    &  & joints & 0.607 \\
	&  & collective & 0.609 \\
	\hline
	\multirow{5}{*}{EPPS} & \multirow{5}{*}{joint} 
	& Joint & 0.365 \\
	& & jointly & 0.497 \\
	& & join- & 0.606 \\
	& & common & 0.607 \\
	& & collective & 0.626 \\
	& & artic- & 0.633\\
	\hline
	\multirow{5}{*}{ECB} & \multirow{5}{*}{joint}
	& Joint & 0.423 \\
	& & jointly & 0.467 \\
	& & common & 0.601 \\
	& & together & 0.632 \\
	& & artic- & 0.663 \\
	& & coordinated & 0.668 \\
\end{tabular}
\caption{Corpora}
\label{tab:Polysemy}
\end{table}
